Plant-pollintor networks worldwide are composed by the same specific building

### 2021-10-08 Draft/Draft.Rmd
Intro ready.

Ecological processes leave distinct structural imprints on species interactions shaping the topology of mutualistic networks. Detecting those relationships is not trivial since they go beyond pair-wise interactions, but may get blurred when considering full network descriptors. However, recent work has shown the network meso-scale can capture this important information. The meso-scale describes network subgraphs representing patterns of interactions between a small number of species (i.e. motifs) and those constitute the building blocks of the whole network. Here, we have compiled 60 networks from 18 different studies and show that some motifs are consitently under- and over-represented worldwide, suggesting that the building blocks of plant-pollintor networks are not random and are associated depending on the motif position and trophic level to specific fucntional groups. For instance, bees are the predominant group found on all motif positions but for plants there are three functional groups associated with corss-pollination that are similarly found on all motif positions. Hence, the distribution of funtional groups on the different positions is not random either which determine in last instance different probabilities of the overall functional group combinations of motifs.

@@

Because of the tremendous complexity of ecological networks, species interactions are generally condensed in metrics that summarize this information [@guimaraes2020]. Plant-pollinator interactions are no exception, and they are often studied with indices that aggregate the information at the species (micro-scale) or network level (macro-scale). Plant-pollinator network level approcahes have proven common invariant structural properties across them including a degree distribution that decays as a power law [@jordano1987], nestedness [@bascompte2003], or modularity in large networks [@olesen2007]. In addition, species phenological overlap, morphological matching and species abundances have been shown to be determinant for the understanding of pairwise plant-pollinator interactions [@bartomeus2016; @stang2006; @peralta2020]. Despite the unquestionable progress with the use of these metrics at both species and network level, condensing the information into a single metric implies the loss of relevant ecological information that obscure the understanding of the ecological processes [@cirtwill2018; @simmons2019].

@@

Sugerencia Alfonso (next paragraph, desde inicio a "(e.g., nestedness)"):Quizás esto lo intentaría sintetizar y meter al final del párrafo anterior y ya empezar en el segundo párrafo a hablar de motifs, como marco de trabajo/nivel de análisis complementario a los anteriores.
---

Traditionally, plant-pollinator research has focused on direct interactions but overlooked indirect interactions (i.e., the mediated effect between two species by a third species), such as facilitative or competitive interactions between plants for pollinators [@moeller2004; @sargent2008]. Despite the widespread nature of indirect interactions in ecological communities [@strauss1991], plant-pollinator research often fails to finely capture those indirect interactions with the conventional analytical tools that condense the information either by species (e.g., interaction frequency) or in single topological indices (e.g., nestedness). Nonetheless, the emerging framework of network motifs (small networks recorded in a given network) in plant-pollinator research, often referred to as the building blocks of a network that depict subsets of interactions [@milo2002], allows to consider both direct and indirect interactions [@simmons2019]. The analysis of motifs in plant-pollinator networks have revealed that the different ecological processes that govern species interactions (e.g., species abundances versus trait-matching) can lead to different patterns of indirect interactions [@simmons2020]. Yet, the global patterns of indirect interactions in real plant-pollinator networks are still unknown (e.g., over- and under- represented motifs).

---
THIS IS A COMMENT
Sugerencia Alfonso: Yo haría un único párrafo hablando del marco de motifs y de sus principales obstáculos: (i) el desconocimiento de los patrones de relaciones indirectas en redes de plantas polinizadores reales y (ii) la desconexión entre las estructuras de los motifs y la ecología de las especies (sus comportamientos, estrategias, etc.). Luego pasaría al párrafo de qué hacemos en este trabajo. 
Digo esto porque la última frase del párrafo anterior me parecía que se quedaba algo colgada.
---

Motifs are abstract representations often decoupled from species ecology. Thus, linking the structural properties of the meso-scale with the species ecology can help the understanding of ecological processes. For instance, different motifs can have different ecological meanings [@simmons2019] and the position within a motif can determine the species functional role [@stouffer2012; @baker2015]. However, it is unclear how the species ecology and life history traits determines the species functional role within the network of interactions [@coux2016]. For example, large pollinators can forage larger distances [@greenleaf2007], deposit greater pollen quantities [@foldesi2021] and handle complex zygomorphic flowers in comparison with small pollinators that are restricted to lower floral complexity [@gong2009]. How this different pollinator behaviors translate into their interaction topology is unknown. Similarly, recent empirical findings indicate that the meso-scale is the best descriptor of plant reproductive success [@allen2021], but little is known on how plants reproductive strategies shape their position within the network of interactions. Although some studies have evaluated plant reproductive strategies in plant-pollinator networks [@tur2013; @lazaro2020], they are often overlooked in a community context [@devaux2014] and rarely incorporated into plant-pollinator network studies. Hence, exploring how the main plant reproductive strategies integrate with the emergent motif framework can shed light on key aspects of ecosystem functioning.

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside functional groups of plants and floral visitors based on functional traits and taxonomic rank, respectively. To obtain plant functional groups, we used a comprehensive dataset that included floral, reproductive and vegetative traits compiled on a larger set of networks. Floral visitor functional groups were divided by the main taxonomical groups that differed in life form and behaviour. Once we split the different plant-pollinator networks into their motif elements, we explored: (i) the overall motif patterns (over and under-represented motifs); (ii) over and under-represented functional groups in the different motif positions; (iii) the probability of finding plant and floral visitor functional groups in the different motif positions; and, (iv) over- and under- represented functional group combinations within the different motifs.

---
# THIS IS A COMMENT (Jose)
# Each os this points (i to iv) corresponds to a figure
# I think that ideally we can have that structure on methods and results (excluding functional groups)
# so far we have (i), (ii) and (iv) on methods
# we lack an explicit mention to (iii) 
# maybe merge it with another research question and just have 3?
# What do you think? 
---

>>>>

Ecological processes leave distinct structural imprints on species interactions shaping the topology of mutualistic networks. Detecting those relationships is not trivial since they go beyond pair-wise interactions, but may get blurred when considering full network descriptors. However, recent work has shown the network meso-scale can capture this important information. The meso-scale describes network subgraphs representing patterns of interactions between a small number of species (i.e. motifs) and those constitute the building blocks of the whole network. Here, we have compiled 60 networks from 18 different studies and show that some motifs are consitently under- and over-represented worldwide, suggesting that the building blocks of plant-pollintor networks are not random. Interestingly, different functional groups are associated to specific motif position across  both trophic levels. HERE I WOULD PUT AN EXAMPLE OF E.G. FAN POSITIONS, DEFINED BY, TEND TO BE OCCUPIED BY BEES AND WHATEVER TYPE OF PLANTS, RESPECTIVELY. This findings open the door to predict how novel interactions will be formed within a community context.

@@

Ecological communities are formed by a plethora of interacting species. Because of the tremendous complexity of this ecological networks, species interactions are generally condensed in metrics that summarize this information [@guimaraes2020]. Plant-pollinator interactions are no exception, and they are often studied with indices that aggregate the information at the species (micro-scale) or network level (macro-scale). For example, plant-pollinator network level approcahes have identified common invariant structural properties across networks, including a degree distribution that decays as a power law [@jordano1987], nestedness [@bascompte2003], or modularity [@olesen2007]. In addition, the local species position within the network can define its degree of specialization [@bluthegXX] or its role connecting the rest of the community [@olesen2007]. Despite the unquestionable progress with the use of these metrics at both species and network level, condensing the information into a single metric implies the loss of relevant ecological information that obscure the understanding of the ecological processes [@cirtwill2018; @simmons2019].

@@

Sugerencia Alfonso (next paragraph, desde inicio a "(e.g., nestedness)"):Quizás esto lo intentaría sintetizar y meter al final del párrafo anterior y ya empezar en el segundo párrafo a hablar de motifs, como marco de trabajo/nivel de análisis complementario a los anteriores.

He modificado un poco el texto, yo creo que se lee bien.
---

Traditionally, plant-pollinator research has focused on direct interactions but overlooked indirect interactions (i.e., the mediated effect between two species by a third species), such as facilitative or competitive interactions between plants for pollinators [@moeller2004; @sargent2008]. Despite the widespread nature of indirect interactions in ecological communities [@strauss1991], plant-pollinator research often fails to finely capture those indirect interactions with the conventional analytical tools that condense the information either by species (e.g., interaction frequency) or in global topological indices (e.g., nestedness). Nonetheless, the emerging framework of network motifs (small sub-networks recorded within a given network) in plant-pollinator research allows to consider both direct and indirect interactions [@simmons2019]. Motifs are often referred to as the building blocks of a network that depict subsets of interactions [@milo2002]. The analysis of motifs in plant-pollinator networks have revealed that the different ecological processes that govern species interactions (e.g., species abundances versus trait-matching) can lead to different patterns of indirect interactions [@simmons2020]. Yet, the global patterns of indirect interactions in real plant-pollinator networks are still unknown (e.g., over- and under- represented motifs).

---
THIS IS A COMMENT
Sugerencia Alfonso: Yo haría un único párrafo hablando del marco de motifs y de sus principales obstáculos: (i) el desconocimiento de los patrones de relaciones indirectas en redes de plantas polinizadores reales y (ii) la desconexión entre las estructuras de los motifs y la ecología de las especies (sus comportamientos, estrategias, etc.). Luego pasaría al párrafo de qué hacemos en este trabajo. 
Digo esto porque la última frase del párrafo anterior me parecía que se quedaba algo colgada.+

Yo para esta version lo veo bien. Se puede mejorar la estructura quizas, pero no se si vale la pena ahora.
---

Despite motifs are abstract representations often decoupled from species ecology, linking the structural properties of the meso-scale with the species ecology can help the understanding of ecological processes. For instance, different motifs can have different ecological meanings [@simmons2019] and the position within a motif can determine the species functional role [@stouffer2012; @baker2015]. However, it is unclear how the species ecology and life history traits determines the species functional role within the network of interactions [@coux2016]. For example, large pollinators can forage larger distances [@greenleaf2007], deposit greater pollen quantities [@foldesi2021] and handle complex zygomorphic flowers in comparison with small pollinators that are restricted to lower floral complexity [@gong2009]. How this different pollinator behaviors translate into their interaction topology is unknown. Similarly, recent empirical findings indicate that the meso-scale is the best descriptor of plant reproductive success [@allen2021], but little is known on how plants reproductive strategies shape their position within the network of interactions. Although some studies have evaluated plant reproductive strategies in plant-pollinator networks [@tur2013; @lazaro2020], this is an  often overlooked aspect in a community context [@devaux2014] and rarely incorporated into plant-pollinator network studies. Hence, exploring how the main plant reproductive strategies integrate with the emergent motif framework can shed light on key aspects of ecosystem functioning.

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside functional groups of plants and floral visitors based on functional traits and taxonomic rank, respectively. To obtain plant functional groups, we used a comprehensive dataset that included floral, reproductive and vegetative traits compiled on a larger set of networks. Floral visitor functional groups were divided by the main taxonomical groups that differed in life form and behaviour. Once we split the different plant-pollinator networks into their motif elements, we explored: (i) if there are common invariant structural properties in the overall motif patterns (i.e. over and under-represented motifs); (ii) which functional groups  are over and under-represented in the different motif positions; and, (iii) if there are over and under-reprsented functional group combinations within a motif.

---
# THIS IS A COMMENT (Jose)
# Each os this points (i to iv) corresponds to a figure
# I think that ideally we can have that structure on methods and results (excluding functional groups)
# so far we have (i), (ii) and (iv) on methods
# we lack an explicit mention to (iii) 
# maybe merge it with another research question and just have 3?
# What do you think? 
#IB: I already unified questions before reading your comment :D
---
